% StudyOS Thesis-Finder — project presentation
% Build: latexmk -pdf studyos-thesis-finder.tex
\documentclass[11pt,aspectratio=169]{beamer}

\usetheme[progressbar=frametitle]{metropolis}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{booktabs}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,shapes.misc}

% Slightly calmer metropolis accents
\definecolor{sosAccent}{HTML}{2A6F77}
\definecolor{sosDark}{HTML}{23373B}
\setbeamercolor{progress bar}{fg=sosAccent}
\setbeamercolor{frametitle}{bg=sosDark}
\setbeamercolor{alerted text}{fg=sosAccent}

\newcommand{\quotebox}[1]{%
  \begin{beamercolorbox}[rounded=true,sep=0.6em]{block body}%
    \itshape #1%
  \end{beamercolorbox}}

\title{StudyOS Thesis-Finder}
\subtitle{From vague interests to a prepared first contact --- with no database to maintain}
\author{StudyOS Team}
\institute{Department of Computer Science, University of Tübingen}
\date{July 2026}

\begin{document}

\maketitle

\begin{frame}{The problem, from both sides}
  \begin{columns}[T]
    \begin{column}{0.48\textwidth}
      \textbf{Students}
      \begin{itemize}
        \item Spend weeks figuring out which chairs even work on their interests
        \item See 4--5 listed topics, assume that is the whole lab, and never reach out
        \item Send generic cold emails to the wrong chairs
      \end{itemize}
    \end{column}
    \begin{column}{0.48\textwidth}
      \textbf{Professors}
      \begin{itemize}
        \item Receive far more inquiries than they can supervise \\ (one chair: 50+ inquiries for 20 slots)
        \item Spend time writing rejections they regret
        \item Get low-signal first contacts that miss the chair's actual research
      \end{itemize}
    \end{column}
  \end{columns}
  \vspace{1em}
  \begin{center}
    \alert{Both sides lose to the same failure: a broken cold start.}
  \end{center}
\end{frame}

\begin{frame}{What 27 professors told us --- before we built anything}
  We interviewed 27 professors in this department. The verdict on a classic
  \emph{topic-listing platform} was blunt:
  \begin{itemize}
    \item \textbf{$\sim$50\%} do not want a central platform --- either already
      overrun with applicants, or convinced topics must be
      \emph{co-created in conversation}, not browsed from a list
    \item \textbf{$\sim$35\%} would try a tool --- but only if zero-friction,
      with no reminder emails, and stable for years
    \item \textbf{$\sim$15\%} recruit through other channels entirely
  \end{itemize}
  \quotebox{``Profs never update the topics. We've had many attempts in the
  department already.''}

  A topic platform was tried here around 2022 and failed --- for
  \alert{incentive} reasons, not UI reasons.
\end{frame}

\begin{frame}{The plan we started with --- and why we abandoned it}
  \begin{columns}[T]
    \begin{column}{0.5\textwidth}
      \textbf{First attempt}\\[0.3em]
      A hosted web app --- FastAPI $\cdot$ Celery $\cdot$ Postgres+pgvector
      $\cdot$ React --- scraping professors and serving a match portal.
      \smallskip
      \begin{itemize}
        \item Covered \textbf{CS Tübingen only}, would have gone stale within months
        \item Assumed HiWis would keep a scraping pipeline alive ``forever'', but with low effort
        \item But still needed someone to maintain the database
      \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
      \textbf{The pivot}\\[0.3em]
      \begin{center}
        \alert{\large A database is the wrong asset.}
      \end{center}
      \smallskip
      The professors' own signal pointed the way: \emph{build for students,
      read what already exists, ask chairs for nothing.}
      \smallskip

      So we stopped maintaining data and started packaging \emph{how to
      search} --- a portable agent skill, live web as the only data source.
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{The idea: ship the method, not the data}
  \begin{center}
    A tool that requires \alert{zero professor participation} and has
    \alert{no database to keep fresh}.
  \end{center}
  Setup is copy-and-paste: drop the skill folder into any standard AI agent
  (Claude, ChatGPT, Gemini) and type \texttt{thesis-finder} --- no install, no
  account. From there the skill \alert{takes the student by the hand}:
  \begin{enumerate}
    \item \textbf{Formulate} --- a guided interview turns vague interests into
      a structured six-dimension profile
    \item \textbf{Discover} --- live search surfaces which chairs, groups, and
      companies actually work on relevant problems
    \item \textbf{Prepare} --- draft a specific, high-signal first contact
  \end{enumerate}
  \smallskip
  The deliverable is a \alert{map of the possibilities} --- ``\emph{now I know
  what exists and where to start.}''
\end{frame}

\begin{frame}{How it works: one entry point, two tracks}
  \begin{center}
  \begin{tikzpicture}[
      stage/.style={draw=sosDark, thick, rounded corners=2pt, fill=sosDark!6,
        align=center, font=\footnotesize, inner sep=4pt, minimum width=24mm},
      track/.style={draw=sosAccent, thick, rounded corners=2pt, fill=sosAccent!8,
        align=center, font=\footnotesize, inner sep=4pt, minimum width=30mm},
      arr/.style={-{Stealth[length=2.5mm]}, thick, sosAccent}]
    \node[stage] (a) {raw input\\+ interview};
    \node[stage, right=6mm of a] (b) {6-dimension\\profile};
    \node[track, right=8mm of b, yshift=7mm] (u) {University track\\all Tübingen faculties};
    \node[track, right=8mm of b, yshift=-7mm] (c) {Industry track\\BW companies};
    \node[stage, right=7mm of u, yshift=-7mm] (o) {map of options\\+ contact draft};
    \draw[arr] (a) -- (b);
    \draw[arr] (b) -- (u);
    \draw[arr] (b) -- (c);
    \draw[arr] (u) -- (o);
    \draw[arr] (c) -- (o);
  \end{tikzpicture}
  \end{center}
  \begin{itemize}
    \item Input as thin as \emph{``I like deep learning and robotics''} ---
      the agent interviews one question at a time to fill the gaps
    \item Profile dimensions: \emph{interests $\cdot$ methods $\cdot$ domain
      $\cdot$ thesis style $\cdot$ skills $\cdot$ no-gos}
    \item Every option is grouped by interest, carries a fit rationale and
      dated evidence, and ends with an honest coverage caveat
  \end{itemize}
\end{frame}

\begin{frame}{The core IP: a search method, not a store}
  \begin{columns}[T]
    \begin{column}{0.46\textwidth}
      \begin{center}
      \begin{tikzpicture}[
          node distance=4mm,
          n/.style={draw=sosDark, thick, rounded corners=2pt, fill=sosDark!6,
            font=\small, inner sep=4pt, minimum width=34mm, align=center},
          arr/.style={-{Stealth[length=2.5mm]}, thick, sosAccent}]
        \node[n] (p) {\textbf{Pass 1}\\filter faculty backbone\\(official listing pages)};
        \node[n, below=of p] (r) {\textbf{Pass 2}\\live web enrichment\\per chair / group};
        \node[n, below=of r] (w) {map of options\\grouped by interest};
        \draw[arr] (p) -- (r);
        \draw[arr] (r) -- (w);
      \end{tikzpicture}
      \end{center}
    \end{column}
    \begin{column}{0.52\textwidth}
      \small
      The value over ``just ask Claude'' lives in two curated references the
      skill carries:
      \begin{itemize}
        \item \textbf{Faculty backbone} --- the official
          \texttt{uni-tuebingen.de} listing pages per faculty, crawled first
          so discovery starts from the real org chart
        \item \textbf{Search strategy} --- the reusable mapping from profile
          dimensions to precise queries, plus filters, dedup, and routing
      \end{itemize}
      A parallel $\sim$100-company BW backbone anchors the industry track.
      Everything is plain, reviewable Markdown. \alert{No runtime database.}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Why no database}
  \renewcommand{\arraystretch}{1.3}
  \small
  \begin{center}
  \begin{tabular}{@{}p{0.46\textwidth}p{0.46\textwidth}@{}}
    \toprule
    \textbf{Database approach (abandoned)} & \textbf{Database-less skill (this project)} \\
    \midrule
    Goes stale within months               & Always current --- reads the live web \\
    CS Tübingen only                       & All faculties, immediately \\
    Needs maintenance                      & Zero ongoing maintenance \\
    Only covers what is curated            & Covers what is publicly visible \\
    Does not scale well to other faculties & Companies use the same principle \\
    \bottomrule
  \end{tabular}
  \end{center}
  \vspace{0.8em}
  \begin{center}
    \alert{The skill is the intelligence; the data comes from the world.}\\

  \end{center}
\end{frame}

\begin{frame}{Privacy and governance}
  \begin{itemize}
    \item \textbf{Student data never leaves the session.} The profile
      (interests, coursework, constraints) exists only in the student's own
      agent conversation --- nothing is stored or transmitted by us.
    \item \textbf{Only public academic data is used}: names, roles, research
      areas, and publications as published by chairs themselves ---
      GDPR-conscious by design.
    \item \textbf{Everything is inspectable.} Skills and references are plain
      text in a public repository; any chair can review --- or ask us to
      correct --- what is said about their group.
    \item \textbf{No fabricated numbers.} A hard project rule: no placeholder
      metrics, no invented team sizes or citation counts.
  \end{itemize}
\end{frame}

\begin{frame}{Where we are, and what comes next}
  \textbf{Behind us}
  \begin{itemize}
    \item Talked to professors and understood what actually goes wrong
    \item Dropped the database idea; built the skill-based tool instead
    \item Made it cover every Tübingen faculty --- and companies as a second track
  \end{itemize}
  \smallskip
  \textbf{Right now}
  \begin{itemize}
    \item Making the tool genuinely \emph{good}, not just working: refining how
      it searches and improving our \alert{evaluation pipeline} that measures
      how well it finds the right chairs
  \end{itemize}
  \smallskip
  \textbf{Still to come}
  \begin{itemize}
    \item A hands-on test with a handful of real students
    \item Getting it into students' hands --- starting with the Fachschaft, and friends
    \item Any suggestions for that?
  \end{itemize}
\end{frame}

\begin{frame}{What we ask the professors (very little)}
  \begin{enumerate}
    \item \textbf{Nothing, by default.} The tool works from your public pages
      and publications as they are.
    \item \textbf{Optionally:} a few research-area keywords, or a plain
      ``looking for / not looking for'' note.
    \item \textbf{Feedback:} if the inquiries you receive start arriving
      better-prepared, tell us --- that is our success metric.
  \end{enumerate}
  \vspace{0.6em}
  \begin{center}
    \alert{Students get better discovery. You get fewer but better inquiries.\\
    Nobody is forced to maintain anything.}
  \end{center}
\end{frame}

\begin{frame}[standout]
  Thank you --- questions?
\end{frame}

\end{document}
